---
title: "P8105_EZ2384_HW3"
output: github_document
editor_options: 
  chunk_output_type: console
---

```{r load_data, message=FALSE}
# Load the required dataset
  library(p8105.datasets)
  library(dplyr)
  library(ggplot2)
  library(tidyverse)
  library(patchwork)
  
  data("ny_noaa")
```

### Problem 1

  - Dataset `ny_noaa` is of size `r nrow(ny_noaa)` x  `r ncol(ny_noaa)`, with variables
      - **id**    [char]: Weather station ID
      - **date**  [date]: Date of observation
      - **prcp**  [int] : Precipitation (tenths of mm)
      - **snow**  [int] : Snowfall (mm)
      - **snwd**  [int] : Snow depth (mm)
      - **tmax**  [char]: Maximum temperature (tenths of degrees C)
      - **tmin**  [char]: Minimum temperature (tenths of degrees C)
  - Missing Data
      - `tmin` and `tmax` contains a great deal of missing values
      - `prcp` and `snow` contains some missing values
      - `id` and `date` are complete

  - Data Cleaning Procedures
      - preliminary cleaning (give reasonable column names, rename column names to be more straightforward)
      - created new columns year,month,day 
      - dropped all NA values 
      - changed all column types except for id and date into integer type to facilitate future conversions
      - changed column values for temperature, precipitation, snowfall so that are given in integer units
      
```{r Data_Cleaning, echo=FALSE}

  snow_data = ny_noaa %>% janitor::clean_names() %>% 
              rename("precipitation(mm)"=prcp, "snowfall(mm)"=snow, "snow_depth(mm)"=snwd, 
                     "maximum_temp(°C)"=tmax, "minimum_temp(°C)"=tmin) %>%
              mutate(year=format(date, "%Y"), month=format(date, "%m"), day=format(date, "%d")) %>%
              mutate(across(`precipitation(mm)`:day, as.integer)) %>%
              mutate(`precipitation(mm)`=`precipitation(mm)`/10,    # use the escape character `` for columns containing special chars
                     `maximum_temp(°C)`=`maximum_temp(°C)`/10,
                     `minimum_temp(°C)`=`minimum_temp(°C)`/10
                     ) %>%   
              drop_na()

  snow_data %>% group_by(`snowfall(mm)`) %>% count(name="count") %>% arrange(desc(count))

```

  - For snowfall, the most commonly observed values are 0, which has a total occurrence of 1112758
  
### Problem 2

```{r}
# Load, tidy, merge, and organize datasets. Exclusion of participants less than 21 yrs and those with missing demographic data.
  demographics =  read.csv("../../Datasets/participant_demographics.csv", skip=4) %>% janitor::clean_names() %>% filter(age>=21) %>%
                  mutate(across(seqn:education, as.character)) %>%  
                  mutate(sex=recode(sex, "1"="male", "2"="female"), 
                         education=recode(education, "1"="below High School", "2"="equivalent to High School", "3"="above High School")) %>% rename(id=seqn) %>% drop_na()

  accelerometers= read.csv("../../Datasets/accelerometers.csv") %>% janitor::clean_names() %>% rename(id=seqn) %>% mutate(id=as.character(id))
  
  merged = left_join(demographics, accelerometers, "id")
  
# Table
  table_by_education =  demographics %>% group_by(education, sex) %>% count(name="count")
  table_by_education

# Plot
  barplot = ggplot(demographics, aes(x = sex, fill = education)) +
            geom_bar(position = "dodge") +
            labs(title = "Distribution of Sex by Education Level", x = "Sex", y = "Count", fill = "Education Level") +
            theme_light() + theme(plot.title=element_text(hjust=0.5))
  barplot

# Aggregated Plot
  aggregated = merged %>% mutate(total_minutes=rowSums(select(.,starts_with("min")))) %>%  relocate("id", "sex", "age", "bmi", "education", "total_minutes")
  
# Plot Total_Minutes against Age
  scatterplot_total = aggregated %>% mutate(age=as.integer(age)) %>%
                      ggplot(., mapping=aes(x=age, y=total_minutes, color=sex)) + geom_point(na.rm=TRUE, size=1) + theme_light() +
                      ggtitle("Scatterplot of Age against Total Minutes") + xlab("Age") + ylab("Total Minutes")+
                      scale_x_continuous(expand=c(0,0), limits=c(0,100)) + theme(plot.title = element_text(hjust=0.5)) + geom_smooth() +
                      facet_grid( . ~ education )
  scatterplot_total
  
# Plot of 24H-Minutes against Median of Acceleratorometers_Readings
  temp_longer = merged %>% pivot_longer(., cols = starts_with("min"), names_to = "minutes", values_to = "readings") %>%
                mutate(minutes = as.numeric(gsub("min","",minutes))) %>% arrange(minutes, education)

  median_min  = temp_longer %>% group_by(minutes) %>% summarise(median_min = median(readings, na.rm=TRUE)) %>%
                mutate(median_min=round(median_min, digits=4)) %>% arrange(minutes)
  
  scatterplot_minutes = ggplot(median_min, mapping=aes(x=minutes, y=median_min)) + geom_point(na.rm=TRUE, size=1) + theme_light() +
                        scale_x_continuous(breaks=seq(0,1400, by=200)) + scale_y_continuous(breaks=seq(0,15,by=2))
                        
```

  - Remark on the barplot showing the distribution of sex by education level
    - Females have higher counts of education level above high school
    - Males have an education level more evenly distributed

  - Remark on the scatterplot of total_activities against age, distinguished by sex, paneled by education level
    - Overall, participants have lower total minutes as age increases, regardless of their sex and education levels.
    - Overall, males tend to have higher total minutes compared to females, regardless of their age and education levels.
    - For those with education level above high school, the total minutes for subjects is relatively smooth and even.
      For those with education level equivalent to high school, the total minutes for subjects has a peak around middle ages.
      For those with education level lower than high school, the total minutes for subjects drops for all age intervals as age increases.
    
  - Remark on the scatterplot of 24H Minutes against Median_Minutes
    - We can see that the median minutes tend to fluctuate greatly during a course of a day, but reaches the lowest point at around 200-300 minutes
      (about 3am to 5 am) and arrived at the peak at about 600-850 minutes(about 10am to 2pm).
    - The curve increases drastically over the course of 5am to 10am; and decreases drastically over the course of 2pm and after.

### Problem 3

- Import, clean, and tidy these data, and describe the resulting dataset.
```{r}
# CityBike Datasets
  city_bike1 = read.csv("../../Datasets/CityBike/CityBike_Jan2020.csv")   %>% janitor::clean_names() %>% mutate("year"="2020", "month"="Jan")
  city_bike2 = read.csv("../../Datasets/CityBike/CityBike_July2020.csv")  %>% janitor::clean_names() %>% mutate("year"="2020", "month"="July")
  city_bike3 = read.csv("../../Datasets/CityBike/CityBike_Jan2024.csv")   %>% janitor::clean_names() %>% mutate("year"="2024", "month"="Jan")
  city_bike4 = read.csv("../../Datasets/CityBike/CityBike_July2024.csv")  %>% janitor::clean_names() %>% mutate("year"="2024", "month"="July")

# Combined CityBike Dataframe
  city_bike_binded = bind_rows(city_bike1,city_bike2,city_bike3,city_bike4) %>%
                     relocate("ride_id", "member_casual", "year", "month", "weekdays", "rideable_type", "start_station_name", "end_station_name") %>%
                     rename("id"="ride_id", "type"="member_casual", "bike"="rideable_type", "start_station"="start_station_name", "end_station"="end_station_name") %>%
                     mutate(weekdays=recode(weekdays, "Monday"="Mon","Tuesday"="Tue", "Wednesday"="Wed", "Thursday"="Thu", "Friday"="Fri", "Saturday"="Sat", "Sunday"="Sun")) %>%
                     mutate(weekdays=factor(weekdays, levels = c("Mon", "Tue", "Wed", "Thu", "Fri", "Sat", "Sun")))

# Table showing total number of rides in each combination of year and month
  table_by_YearMonthType =  city_bike_binded %>% group_by(year, month, type) %>% count(name="count")
  table_by_YearMonthType

# Top 5 originating stations and the corresponding number of rides in July.2024
  top5_starting_positions = city_bike4 %>% group_by(start_station_name) %>% count(name="count")
  head(top5_starting_positions, n=5)
  
# Median Riding Duration depending on Covariates
  median_duration_year     =  city_bike_binded %>% group_by(year) %>% summarise(median_duration = median(duration, na.rm = TRUE))
  median_duration_month    =  city_bike_binded %>% group_by(month) %>% summarise(median_duration = median(duration, na.rm = TRUE))
  median_duration_weekdays =  city_bike_binded %>% group_by(weekdays) %>% summarise(median_duration = median(duration, na.rm = TRUE))
  median_duration_membership  =  city_bike_binded %>% group_by(type) %>% summarise(median_duration = median(duration, na.rm = TRUE))
  median_duration_bike        =  city_bike_binded %>% group_by(bike) %>% summarise(median_duration = median(duration, na.rm = TRUE))

# Subplots of Median Riding Duration depending on year/month/weekday
  subplot1 = ggplot(median_duration_year, aes(x = year, y = median_duration)) +
             geom_bar(stat="identity", fill="steelblue", width=0.75) +
             xlab("Year") + ylab("Median Duration") + ggtitle("Median Duration(Year)") +
             theme_light() + theme(plot.title = element_text(hjust=0.5))
  
  subplot2 = ggplot(median_duration_month, aes(x=month, y=median_duration)) +
             geom_bar(stat="identity", fill="darkblue", width=0.75) +
             xlab("Month") + ylab("Median Duration") + ggtitle("Median Duration(Month)") +
             theme_light() + theme(plot.title = element_text(hjust=0.5))
  
  subplot3 = ggplot(median_duration_weekdays, aes(x=weekdays, y=median_duration)) +
             geom_bar(stat="identity", fill="royalblue") +
             xlab("Weekday") + ylab("Median Duration") + ggtitle("Median Duration(Weekday)") +
             theme_light() + theme(plot.title = element_text(hjust=0.5))

# Final plot of Median Riding Duration depending on Year/Month/Weekday
  subplot1 + subplot2 + subplot3

# Figure showing impact of month, membership status, and bike type on the distribution of ride duration

  subplot4 = ggplot(median_duration_membership, aes(x = type, y = median_duration)) +
             geom_bar(stat="identity", fill="darkgreen", width=0.75) +
             xlab("Membership Type") + ylab("Median Duration") + ggtitle("Median Duration(Membership Type)") +
             theme_light() + theme(plot.title = element_text(hjust=0.5))
  
  subplot5 = ggplot(median_duration_bike, aes(x=bike, y=median_duration)) +
             geom_bar(stat="identity", fill="darkred", width=0.75) +
             xlab("Bike Type") + ylab("Median Duration") + ggtitle("Median Duration(Bike Type)") +
             theme_light() + theme(plot.title = element_text(hjust=0.5))

  # Final Plot of Median Riding Duration depending on Membership_Type/Bike_Type
  subplot4 + subplot5
  
```

- Description of CityBike Datasets
  - CityBike Datasets consists of 4 component datasets with each consisting of 7 variables:
    - ride_id : unique identifier for CityBike user for the particular trip
    - rideable_type: type of bike used by user for the particular trip
    - weekdays: weekday info when riding for the particular trip
    - duration: duration or ride for the particular trip
    - start_station_name: set-off station name for the particular trip
    - end_station_name: ending station name for the particular trip
    - member_casual: membership type of the user for the particular trip
  - Each dataset has dimensions as below:
    - Dataset1: info of Jan.2020, dimension=`r nrow(city_bike1)` x 7
    - Dataset2: info of July.2020, dimension=`r nrow(city_bike2)` x 7
    - Dataset3: info of Jan.2024, dimension=`r nrow(city_bike3)` x 7
    - Dataset4: info of July.2024, dimension=`r nrow(city_bike4)` x 7
  - Completeness:
    - Each component dataset is complete and without missing values

- Comment on table_by_YearMonthType
  - The total number of riders rise considerably each year from 2020 to 2024, and has risen from 12480 riders in Jan.2020 to 47156 riders in July.2024.
  - The total number of members rise steadily and considerably each time interval from Jan.2020 to July.2020 to Jan.2024 to July.2024
  - Overall, for each timeslot, the number of members are considerably lower compared to the number of casual riders at that time.

- Remark on plot showing median durations depending on year/month/weekday
  - We can see a clear pattern that, citybike users in weekends are considerably more active compared to them in weekdays.
  - The median users in 2024 is lower than that in 2020; and the median users in July is higher than that in January.
    However, without more year groups and/or more month groups, we cannot conclude for sure whether there is a true pattern or the observed difference is due merely to chance.
  
- Remark on plot showing median durations depending on membership_type/bike_type
  - Overall, those who are casual riders have a longer biking duration compared to members by a great amount.
  - Overall, users spend longer time using classic bikes compared to electric bikes, but this difference is not big.
