---
title: "P8105_EZ2384_HW3"
output: github_document
---

```{r load_data, message=FALSE}
# Load the required dataset
  library(p8105.datasets)
  data("ny_noaa")
```

### Problem 1

  - Dataset `ny_noaa` is of size `r nrow(ny_noaa)` x  `r ncol(ny_noaa)`, with variables
      - **id**    [char]: Weather station ID
      - **date**  [date]: Date of observation
      - **prcp**  [int] : Precipitation (tenths of mm)
      - **snow**  [int] : Snowfall (mm)
      - **snwd**  [int] : Snow depth (mm)
      - **tmax**  [char]: Maximum temperature (tenths of degrees C)
      - **tmin**  [char]: Minimum temperature (tenths of degrees C)
  - Missing Data
      - `tmin` and `tmax` contains a great deal of missing values
      - `prcp` and `snow` contains some missing values
      - `id` and `date` are complete

```{r Data_Cleaning}
  ny_noaa %>% janitor::clean_names()

```


- Do some data cleaning. Create separate variables for year, month, and day. Ensure observations for temperature, precipitation, and snowfall
  are given in reasonable units. For snowfall, what are the most commonly observed values? Why?
- Make a two-panel plot showing the average max temperature in January and in July in each station across years. Is there any observable
  interpretable structure? Any outliers?
- Make a two-panel plot showing (i) tmax vs tmin for the full dataset (note that a scatterplot may not be the best option); and (ii) make a plot
  showing the distribution of snowfall values greater than 0 and less than 100 separately by year.
  
```{r}
ny_noaa
```
  
  