---
title: "P8105_EZ2384_HW3"
output: github_document
editor_options: 
  chunk_output_type: console
---

```{r load_data, message=FALSE}
# Load the required dataset
  library(p8105.datasets)
      data("ny_noaa")
  library(dplyr)
  library(ggplot2)
  library(tidyverse)
```

### Problem 1

  - Dataset `ny_noaa` is of size `r nrow(ny_noaa)` x  `r ncol(ny_noaa)`, with variables
      - **id**    [char]: Weather station ID
      - **date**  [date]: Date of observation
      - **prcp**  [int] : Precipitation (tenths of mm)
      - **snow**  [int] : Snowfall (mm)
      - **snwd**  [int] : Snow depth (mm)
      - **tmax**  [char]: Maximum temperature (tenths of degrees C)
      - **tmin**  [char]: Minimum temperature (tenths of degrees C)
  - Missing Data
      - `tmin` and `tmax` contains a great deal of missing values
      - `prcp` and `snow` contains some missing values
      - `id` and `date` are complete

  - Data Cleaning Procedures
      - preliminary cleaning (give reasonable column names, rename column names to be more straightforward)
      - created new columns year,month,day 
      - dropped all NA values 
      - changed all column types except for id and date into integer type to facilitate future conversions
      - changed column values for temperature, precipitation, snowfall so that are given in integer units
      
```{r Data_Cleaning, echo=FALSE}

  snow_data = ny_noaa %>% janitor::clean_names() %>% 
              rename("precipitation(mm)"=prcp, "snowfall(mm)"=snow, "snow_depth(mm)"=snwd, 
                     "maximum_temp(°C)"=tmax, "minimum_temp(°C)"=tmin) %>%
              mutate(year=format(date, "%Y"), month=format(date, "%m"), day=format(date, "%d")) %>%
              mutate(across(`precipitation(mm)`:day, as.integer)) %>%
              mutate(`precipitation(mm)`=`precipitation(mm)`/10,    # use the escape character `` for columns containing special chars
                     `maximum_temp(°C)`=`maximum_temp(°C)`/10,
                     `minimum_temp(°C)`=`minimum_temp(°C)`/10
                     ) %>%   
              drop_na()

  snow_data %>% group_by(`snowfall(mm)`) %>% count(name="count") %>% arrange(desc(count))

```

  - For snowfall, the most commonly observed values are 0, which has a total occurrence of 1112758
  
### Problem 2

  *  Load, tidy, merge, and otherwise organize the data sets.
     Exclusion of participants less than 21 yrs and those with missing demographic data.
     Encoded dataset so that all columns in the demographics dataset are converted into character type.

```{r}
  demographics =  read.csv("../../Datasets/participant_demographics.csv", skip=4) %>% janitor::clean_names() %>% filter(age>=21) %>%
                  mutate(across(seqn:education, as.character)) %>%  
                  mutate(sex=recode(sex, "1"="male", "2"="female"), 
                         education=recode(education, "1"="lower than High School", "2"="equivalent to High School", "3"="above High School")) %>%                   rename(id=seqn) %>% drop_na()

  accelerometers= read.csv("../../Datasets/accelerometers.csv") %>% janitor::clean_names() %>% rename(id=seqn) %>% mutate(id=as.character(id))
  
  merged = left_join(demographics, accelerometers, "id")
```

  * Produce a reader-friendly table for the number of men and women in each education category
    create a visualization of the age distributions for men and women in each education category
    Comment on these items.
```{r}
  # Table
  table_by_education =  demographics %>% group_by(education, sex) %>% count(name="count")
  table_by_education
  
  # Plot
  barplot = ggplot(demographics, aes(x = sex, fill = education)) +
            geom_bar(position = "dodge") +
            labs(title = "Distribution of Sex by Education Level", x = "Sex", y = "Count", fill = "Education Level") +
            theme_light()
  barplot

```
  
  - From the bar graph we can see that
    - Females have higher counts of education level above high school
    - Males have an education level more evenly distributed
  
  * Using your tidied dataset, aggregate across minutes to create a total activity variable for each participant.
    Plot these total activities(y-axis) against age(x-axis).
    Your plot should compare men to women and have separate panels for each education level.
    Include a trend line or a smooth line to illustrate differences.
    Comment on your plot.
    
```{r}
  # Aggregated Plot
  aggregated = merged %>% mutate(total_minutes=rowSums(select(.,starts_with("min")))) %>% 
                          relocate("id", "sex", "age", "bmi", "education", "total_minutes")
  
  # Plot Total_Minutes against Age
  scatterplot = aggregated %>% mutate(age=as.integer(age)) %>%
                ggplot(., mapping=aes(x=age, y=total_minutes, color=sex)) + geom_point(na.rm=TRUE, size=1) + theme_light() +
                ggtitle("Scatterplot of Age against Total Minutes") + xlab("Age") + ylab("Total Minutes")+
                scale_x_continuous(expand=c(0,0), limits=c(0,100)) + theme(plot.title = element_text(hjust=0.5)) + geom_smooth() +
                facet_grid( . ~ education )
  scatterplot
  
```
    
  - From the scatterplot we can see that
    - Overall, participants have lower total minutes as age increases, regardless of their sex and education levels.
    - Overall, males tend to have higher total minutes compared to females, regardless of their age and education levels.
    - For those with education level above high school, the total minutes for subjects is relatively smooth and even.
      For those with education level equivalent to high school, the total minutes for subjects has a peak around middle ages.
      For those with education level lower than high school, the total minutes for subjects drops for all age intervals as age increases.
    
  * Make a three-panel plot that shows the 24-hour activity time courses for each education level and use color to indicate sex.
    Describe in words any patterns or conclusions you can make based on this graph;
    including smooth trends may help identify differences.

### Problem 3

